### 2015-11-10 section/abstract.tex
finish abstract

Writing efficient software is difficult.
Design and implementation defects can easily
lead to severe performance 
degradation. Techniques that help
diagnose performance problems are desired.
Unfortunately, existing techniques are still preliminary.
%failing to provide the desired diagnosis coverage, accuracy, and performance. 
Performance-bug detectors can identify specific type
of inefficient computation, but are not suitable for accurately diagnosing
a wide variety of real-world performance problems. 
Profilers can locate code regions that consume resources, but 
not the ones that \textit{waste} resources.
Statistical performance diagnosis can
identify loops or branches that are most correlated with a performance
symptom, but cannot decide whether and why a loop is inefficient, and
how developers might fix it.

In this paper, we first design a root-cause
and fix-strategy taxonomy for real-world inefficient loops, 
one of the most common performance problems in the field.
We then design a static-dynamic hybrid analysis tool, \Tool, to
provide accurate performance diagnosis for loops.
We further use sampling techniques to lower the run-time overhead without
degrading the accuracy or latency of \Tool diagnosis. 
Evaluation using real-world performance
problems shows that \Tool can provide better coverage and accuracy than
existing techniques, with low overhead.