### 2015-11-18 section/tab_root.tex

\def\cca#1{\cellcolor{black!#10}\ifnum #1>5\color{white}\fi{#1}}

\begin{table*}[tb!]
%\begin{adjustwidth}{-.5in}{-.5in}
\small
\centering
{
\begin{tabular}{lccccccccccccc}
\hline
                      & Apache &Chrome & GCC  & Mozilla & MySQL &&M$_{\text{emoization}}$  & B$_{\text{atching}}$   &  C & S  & O$_{\text{ther}}$ && Total\\
\cmidrule{2-6}
\cmidrule{8-12}
                      & 11    & 4     & 8    & 12      & 10     && 12  & 12 & 4  & 9  & 8  && 45\\
\hline
\multicolumn{1}{l}{Cross-{\bf iteration} Redundancy}
&7&1&2&1&1&&\cca{7}&\cca{4}&0&0&\cca{1}&&12\\
\multicolumn{1}{l}{ Cross-{\bf loop} Redundancy}
&3&0&2&2&2&&\cca{5}&\cca{4}&\cca{0}&\cca{0}&\cca{0}&&9\\
\multicolumn{1}{l}{ {\bf 0*1?} Resultless}
&0&0&0&2&3&&\cca{0}&\cca{0}&\cca{4}&\cca{1}&\cca{0}&&5\\
\multicolumn{1}{l}{{\bf [0$|$1]*} Resultless}
&0&1&1&1&1&&\cca{0}&\cca{0}&\cca{0}&\cca{4}&\cca{0}&&4\\
\multicolumn{1}{l}{ {\bf 0*} Resultless}
&0&0&0&0&0&&\cca{0}&\cca{0}&\cca{0}&\cca{0}&\cca{0}&&0\\
\multicolumn{1}{l}{{\bf 1*} Resultless}
&1&2&3&6&3&&\cca{0}&\cca{4}&\cca{0}&\cca{4}&\cca{7}&&15\\
  %MySQL15811 is moved from 1* to here; consider it as fixed by M
\hline
\end{tabular}
}
%\end{adjustwidth}
\caption{The distribution among different applications and different fix strategies for bugs with each root-cause category.
(S: conditionally skip the loop; C: change the data structure)
}
\label{tab:root}
\end{table*}