### 2016-06-01 section/7_con.tex
check in the technical report

\section{Conclusion}
\label{sec:con}
Performance diagnosis is time consuming and also critical for 
complicated modern software. \Tool tries to automatically 
pin-point the root cause of
the most common type of real-world performance problems, inefficient loops,
and provide fix-strategy suggestions to developers. It achieves the 
coverage, accuracy, and performance goal of performance diagnosis by leveraging
(1) a comprehensive root-cause taxonomy; (2) a hybrid static-dynamic program
analysis approach; and (3) customized random sampling that is a natural fit for 
performance diagnosis.
Our evaluation shows that \Tool can accurately identify detailed root causes
of real-world inefficient loop problems and provide fix-strategy suggestions. 
Future work can further improve \Tool by providing more detailed fix
suggestions and providing more information to help diagnose and fix
1* loops.

\acks
This work is supported in part by NSF grants IIS-1546543, CNS-1514256, CCF-1439091, and CCF-1514189; 
and Alfred P. Sloan Research Fellowship.