### 2016-08-25 section/0_intro.tex
finish ab and intro

\section{Introduction}
\label{sec:intro}
\subsection{Motivation}

Performance bugs\footnote{We also refer performance bugs as performance problems
following previous works in this area~\cite{PerfBug,Alabama,SongOOPSLA2014}} 
are software implementation mistakes that cause unnecessary performance
degradation in software. They widely exist in deployed software due to the 
complexity of modern software
~\cite{PerfBug,perf.fse10,rily.perftest,perfantipattern,xiao13:context}. 
They annoy end users and waste energy during production runs, and 
have already caused highly publicized failures~\cite{ACA-health,colorado}.
Tools that can help
developers quickly and accurately diagnose performance problems
are sorely desired.

Performance diagnosis is different from performance testing or
bug detection: it does \textit{not} aim to expose previously unknown performance
anomalies. 
Instead, it is similar with general failure diagnosis: 
it is applied \textit{after} an unexpected software behavior (i.e., symptom)
is observed, hoping to identify the root cause of this symptom and suggest
strategies to eliminate this symptom.
In the context of performance problems, the symptom is execution 
slowness \cite{SongOOPSLA2014}; the root cause is about
which code region is inefficient and why.
An effective diagnosis tool can help developers quickly and correctly understand 
and fix already-exposed performance symptoms.

Also like general failure diagnosis, ideal performance diagnosis tools should
satisfy three criteria.
\begin{itemize}
\item Coverage. 
Real-world performance problems are caused by a wide variety of reasons.
A good diagnosis tool should handle a good portion of them.

\item Accuracy. 
Which code regions are inefficient and why they are inefficient
need to be accurately identified.

\item Performance. 
Diagnosis often requires collecting run-time information. The lower the overhead
is, the easier for the tool to deploy, especially for 
production-run usage. 
\end{itemize}

No existing tools can satisfy these
three requirements.

Profiling is the state of practice in performance diagnosis.
It is far from providing the desired \textit{accuracy}, as it is designed to
tell where computation resources are spent, 
but not where and why resources are wasted. 
In many cases, the root-cause function may not even
get ranked by the profiler \cite{SongOOPSLA2014}.

Performance bug detection tools use program analysis to identify
code regions that match specific inefficiency patterns 
\cite{Alabama,CARAMEL, Cachetor,Xu:2010:FLD:1806596.1806617,Dufour:2008:STC:1453101.1453111, Xu:2009:GFP:1542476.1542523, Xu:2010:DIC:1806596.1806616,IsilDillig.PLDI15}. 
Unfortunately, these tools are not designed and hence are unsuitable
to identify code regions that contribute to a specific slowness symptom.
They do not provide \textit{coverage} for a wide
variety of real-world problems; they are not guided by
performance failure symptoms, and hence are at disadvantage in terms of diagnosis
\textit{accuracy}; dynamic detection tools often lead to 
10X slowdowns or more \cite{Cachetor,Xu:2010:FLD:1806596.1806617,Alabama}, 
not ideal in terms of \textit{performance}.

Recently, progress has been made on statistical debugging for
performance diagnosis \cite{SongOOPSLA2014}.
This approach compares runs with and without problematic performance, and
identifies control-flow constructs, such as a branch $b$ or 
a loop $l$, that are most correlated with
the execution slowness.
Unfortunately, this approach is \textit{not} effective
for loop-related 
performance problems, which contribute to two thirds of
real-world performance problems studied in previous work 
\cite{SongOOPSLA2014,PerfBug}. It cannot
tell whether and how loop $l$ is inefficient 
and hence is limited in its diagnosis \textit{coverage} and \textit{accuracy}.

\begin{figure}
\centering
\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/GCC27733.c}}
\caption{A real-world performance bug in GCC (the `-' and `+' demonstrate the patch; variable and 
function names are simplified for demonstration purposes)}
\label{fig:GCC27733}
\end{figure}

Figure~\ref{fig:GCC27733} shows a performance problem in GCC. 
Recursive function \texttt{mult\_alg} computes the best algorithm for 
multiplying \texttt{t}, a time-consuming computation. 
At run time, \texttt{mult\_alg} is often invoked for many times, and often
with the same parameter partly due to its recursive nature.
To avoid redundant computation across different instances of
\texttt{mult\_alg}, developers used a
hash-table
\texttt{alg\_hash} to remember which parameter \texttt{t} has been processed
in the past and what is the result. 
Unfortunately, a mistake in the type declaration of hash-table
entry \texttt{hash\_entry} makes the memoization useless for large \texttt{t}.
In many cases, a slow path is taken, when the fast path should have been taken.
This mistake does not affect software correctness, but causes a large amount
of redundant computation\footnote{This will be considered as a
loop-redundancy problem later in this paper, as we consider
recursive functions as a special case of loops.} 
and hurts GCC performance severely,
causing hundreds of times slow down for GCC test cases.

Debugging this performance problem is challenging. According to 
the discussion forum of GCC, developers identified
\texttt{mult\_alg} as the most time-consuming function
through profilers early on. However, they did not figure out 
whether \texttt{mult\_alg} is inefficient, which part of it is 
inefficient (\texttt{mult\_alg} is a large function with about 400 lines of 
code), and how it is inefficient, until several weeks later.

If a tool can tell developers not only \emph{which} loop or function is
responsible for execution slowness, but also \emph{why} and \emph{how}
it is inefficient, 
diagnosis and fixing
would be much easier.

\iffalse
%to save space
Clearly, more research is needed to improve the state of the art of performance
diagnosis --- better diagnosis
coverage, better diagnosis accuracy, and low run-time
overhead for common performance problems,
especially loop-related performance problems.
\fi

\subsection{Contributions}
This paper presents a tool \Tool that can help effectively diagnose
inefficient loop problems, the most common type of performance problems
\cite{SongOOPSLA2014,PerfBug}, with good coverage, accuracy, and performance.
After users or developers observe a performance failure symptom, \Tool can 
automatically judge whether the symptom is caused by inefficient loops and
provide detailed root-cause information that helps understand and fix 
inefficient loops.

\Tool tackles this challenging problem in three steps.
%}

First, figuring out a root-cause taxonomy for inefficient loops.
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
Our taxonomy categorizes all inefficient loops into two main categories:
\emph{resultless}, when a large amount of computation produces
no side effects, and \emph{redundancy}, when a large amount of 
computation produces already-available results.
Each main category is further divided to sub-categories.
We strive to make the taxonomy both general enough to cover
common inefficient loops, and specific enough to guide the design
of \Tool and eventually help developers understand
and fix performance problems
(Section \ref{sec:study}).

Second, building a tool(kit) \Tool that can automatically and accurately
identify whether and how a suspicious loop is inefficient, 
following these principles (Section \ref{sec:design}):

\begin{itemize}
\item Focused checking. 
Different from performance-bug detectors that blindly check the whole
software, \Tool focuses on loops that are
most correlated with performance symptoms. 
This focus is crucial for \Tool to achieve both high
accuracy and high coverage.

\item Taxonomy guided design. To provide good coverage, we follow
the root-cause taxonomy discussed above and design analysis routines for
every root-cause category. Given a candidate loop, \Tool
applies a series of analysis to see if it matches any type of inefficiency.
%We will also put the analysis results from different analysis routines together
%to figure out the most likely root cause category for an inefficient loop.

\item Static-dynamic hybrid analysis.
Static analysis alone cannot accurately identify 
inefficiency root causes, because some inefficiency problems only
happen or matter under specific workload. However, pure dynamic analysis will 
cause too large 
run-time overhead. Therefore, we use a hybrid approach to achieve both
performance and accuracy goals.
\end{itemize}

Third, using sampling to further lower the run-time overhead of \Tool, without
degrading diagnosis capability. Random sampling is a natural fit for performance
diagnosis due to the repetitive nature of inefficient loops (Section \ref{sec:inst}).

We evaluated \Tool on \allbugs real-world performance problems,
coming from two representative benchmark suites 
\cite{SongOOPSLA2014,Alabama}. 
Evaluation results show that \Tool can accurately identify detailed
root cause for all benchmarks and provide correct fix-strategy suggestion for
most benchmarks. All of these are achieved with low run-time overhead.