### 2016-08-25 section/7_con.tex

\section{Conclusion}
\label{sec:con}
Performance diagnosis is time consuming and also critical for 
complicated modern software. \Tool tries to automatically 
pin-point the root cause of
the most common type of real-world performance problems, inefficient loops,
and suggest fix strategies to developers. It achieves the 
coverage, accuracy, and performance goal by leveraging
(1) a comprehensive root-cause taxonomy; (2) a hybrid program
analysis approach; and (3) customized random sampling that is a natural fit for 
performance diagnosis.
Our evaluation shows that \Tool can accurately identify detailed root causes
of real-world inefficient loops and suggest fix strategies.
Future work can improve \Tool by providing more detailed fix
suggestions and more information to diagnose and fix
1* loops.

%\acks
%This work is supported in part by NSF grants IIS-1546543, CNS-1514256, CCF-1439091, and CCF-1514189; 
%and Alfred P. Sloan Research Fellowship.